---
title: "Codebook"
output: pdf_document
toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Codebook for Getting and Cleaning Data Course Project

This codebook contains information about the original raw data and the final output
that is produced by this project. Please refer to the README.PDF for details on the 
code blocks and how they work together.

## Overview of the HAR Dataset (aka Human Activity Recognition Using Smartphones Data Set)

### Citation for Original Study
Davide Anguita, Alessandro Ghio, Luca Oneto, Xavier Parra and Jorge L. Reyes-Ortiz. A Public Domain Dataset for Human Activity Recognition Using Smartphones. 21th European Symposium on Artificial Neural Networks, Computational Intelligence and Machine Learning, ESANN 2013. Bruges, Belgium 24-26 April 2013.

### Dataset Information on Original Data Set
The original raw data was obtained from http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones  


These are the critical summary data:

**Abstract** Human Activity Recognition database built from the recordings of 30 subjects performing activities of daily living (ADL) while carrying a waist-mounted smartphone with embedded inertial sensors

**Data Set Characteristics:** Multivariate, Time-Series

**Number of Observations:** 10299

**Number of Variables:** 561

### Data Set Abstract on Original Data Set

>The experiments have been carried out with a group of 30 volunteers within an age 
bracket of 19-48 years. Each person performed six activities (WALKING, 
WALKING_UPSTAIRS, WALKING_DOWNSTAIRS, SITTING, STANDING, LAYING) wearing a 
smartphone (Samsung Galaxy S II) on the waist. Using its embedded accelerometer 
and gyroscope, we captured 3-axial linear acceleration and 3-axial angular velocity 
at a constant rate of 50Hz. The experiments have been video-recorded to label the 
data manually. The obtained dataset has been randomly partitioned into two sets, 
where 70% of the volunteers was selected for generating the training data and 30% the test data.

>The sensor signals (accelerometer and gyroscope) were pre-processed by applying 
noise filters and then sampled in fixed-width sliding windows of 2.56 sec and 50% 
overlap (128 readings/window). The sensor acceleration signal, which has gravitational 
and body motion components, was separated using a Butterworth low-pass filter into body 
acceleration and gravity. The gravitational force is assumed to have only low frequency 
components, therefore a filter with 0.3 Hz cutoff frequency was used. From each window, 
a vector of features was obtained by calculating variables from the time and frequency domain.

### Attribute Information on Original Data Set
For each record in the dataset the following is provided:
- Triaxial acceleration from the accelerometer (total acceleration) and the estimated body acceleration.
- Triaxial Angular velocity from the gyroscope.
- A 561-feature vector with time and frequency domain variables.
- An activity label.
- An identifier of the subject who carried out the experiment.


### Data File Location for Original Data Set

The original .zip file with data can be found at https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip  

### File Structure for Original Data Set

The authors original variable explanations are reproduced here for your convenience.

#### Original Variable Descriptions
```{r comment=""}
cat(readLines("UCI HAR Dataset/features_info.txt"), sep = "\n")
```

#### Original Variable List
```{r comment=""}
cat(readLines("UCI HAR Dataset/features.txt"), sep = "\n")
```

## Overview of the cleanData Data Set

These are the critical summary data:

**Number of Observations:** 10299

**Number of Variables:** 88  


The cleanData data set contains the the following attributes that were done in sequence:

1. Merges the training and the test sets to create one data set.

2. Extracts only the measurements on the mean and standard deviation for each measurement. 

3. Uses descriptive activity names to name the activities in the data set

4. Appropriately labels the data set with descriptive variable names. 

*See README.PDF for the code that implemented these steps*

### File Structure for cleanDataset

Variable information for the cleanDataset is included here for convenience.
```{r comment=""}
cat(readLines("cleanDataStructure.txt"), sep = "\n")
```

## Overview of the avgCleanDataset Data Set

These are the critical summary data:

**Number of Observations:** 180

**Number of Variables:** 88  

The avgCleanDataset represents the average of each variable for each activity and 
each subject. It is derived from the cleanDataset mentioned above. 

*See README.PDF for the code that implemented these steps*

Variable information for the avgCleanDataset is included here for convenience.
```{r comment=""}
cat(readLines("avgCleanDatasetStructure.txt"), sep = "\n")
```